A Speculative Approach to Parallelization in Particle Swarm Optimization

### 2011-11-12 masters_thesis/si_paper/paper.tex
Updated with very minor revisions

\begin{table}
  \caption{Summary of results for function Sphere with 20 dimensions,
  measuring number of time steps to reach a value of 1e-35}
  \label{tab:sphere-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&0\%&N/A&N/A\\
  \hline
  Standard Complete&100\%&931.0&19.1\\
  \hline
  SEPSO Ring&0\%&N/A&N/A\\
  \hline
  SEPSO Random&85\%&972.8&15.8\\
  \hline
  Pick Best Ring&100\%&917.4&21.0\\
  \hline
  Pick Best Random&100\%&768.1&14.3\\
  \hline
  Pick Best Pruned Ring&100\%&967.8&9.9\\
  \hline
  Pick Best Pruned Random&100\%&693.5&10.4\\
  \hline
  \textbf{Pick Best Pruned Complete}&\textbf{100\%}&\textbf{389.6}&
    \textbf{9.6}\\
  \hline
  Social Promotion Pruned Ring&0\%&N/A&N/A\\
  \hline
  Social Promotion Pruned Random&100\%&971.1&10.6\\
  \hline
  Social Promotion Pruned Complete&100\%&777.8&16.7\\
  \hline
  Many Iterations Ring&100\%&575.2&9.8\\
  \hline
  Many Iterations Random&100\%&442.4&7.1\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{Summary of results for function Schwefel with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:schwefel-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&0\%&N/A&N/A\\
  \hline
  Standard Complete&100\%&837.2&44.8\\
  \hline
  SEPSO Random&0\%&N/A&N/A\\
  \hline
  Pick Best Random&5\%&938.0&0.0\\
  \hline
  \textbf{Pick Best Pruned Random}&\textbf{100\%}&\textbf{639.9}&\textbf{29.4}\\
  \hline
  \textbf{Pick Best Pruned Complete}&\textbf{100\%}&\textbf{597.7}&
	\textbf{77.9}\\
  \hline
  Social Promotion Pruned Random&75\%&938.0&33.0\\
  \hline
  Social Promotion Pruned Complete&100\%&815.2&43.5\\
  \hline
  Many Iterations Random&100\%&783.5&94.5\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{Summary of results for function Rastrigin with 20 dimensions,
  measuring number of time steps to reach a value of 20}
  \label{tab:rastrigin-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&25\%&743.4&212.0\\
  \hline
  Standard Random&100\%&372.1&86.6\\
  \hline
  \textbf{Standard Complete}&\textbf{95\%}&\textbf{273.1}&\textbf{99.7}\\
  \hline
  SEPSO Ring&5\%&471.0&0.0\\
  \hline
  SEPSO Random&15\%&208.0&17.0\\
  \hline
  Pick Best Ring&35\%&489.0&152.0\\
  \hline
  Pick Best Random&15\%&136.0&40.1\\
  \hline
  Pick Best Pruned Ring&95\%&545.9&208.5\\
  \hline
  Pick Best Pruned Random&10\%&95.0&11.0\\
  \hline
  Pick Best Pruned Complete&5\%&59.0&0.0\\
  \hline
  Social Promotion Pruned Ring&5\%&580.0&0.0\\
  \hline
  Social Promotion Pruned Random&80\%&278.2&136.3\\
  \hline
  Social Promotion Pruned Complete&50\%&200.5&130.2\\
  \hline
  Many Iterations Ring&10\%&117.0&25.0\\
  \hline
  Many Iterations Random&5\%&57.0&0.0\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank6}
  \caption{Function Griewank with 20 dimensions.  Each method performs one
  evaluation on each of 800 processors per time step.  Instead of showing
  average function value, we show the percentage of runs that have found the
  global optimum by each iteration.  All algorithms use the Ring topology.}
  \label{fig:basic-griewank1}
\end{figure}

\begin{table}
  \caption{Summary of results for function Griewank with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:griewank-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&762.2&19.4\\
  \hline
  SEPSO Ring&100\%&426.0&81.0\\
  \hline
  \textbf{Pick Best Ring}&\textbf{100\%}&\textbf{272.0}&\textbf{17.7}\\
  \hline
  \textbf{Pick Best Pruned Ring}&\textbf{100\%}&\textbf{282.9}&\textbf{10.7}\\
  \hline
  Social Promotion Pruned Ring&100\%&482.0&16.6\\
  \hline
  Many Iterations Ring&84\%&183.2&49.4\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank1}
  \caption{Function Griewank with 20 dimensions.  Each method performs one
  evaluation of the objective function on each of 240 processors per time step.
  All algorithms use the Ring topology.}
  \label{fig:basic-griewank2}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank3}
  \caption{Function Griewank with 20 dimensions, comparing Pick Best with
  results from \fig{basic-griewank2}.  Each method performs one evaluation on
  each of 240 processors per time step.  All algorithms use the Ring topology.}
  \label{fig:griewank-pickbest}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank4}
  \caption{Function Griewank with 20 dimensions, comparing pruning with the
  best results from \fig{griewank-pickbest}.  Each method performs one
  evaluation on each of 240 processors per time step.  All algorithms use the
  Ring topology.}
  \label{fig:griewank-pruned}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank5}
  \caption{Function Griewank with 20 dimensions, comparing speculating several
  iterations ahead with selections from \fig{griewank-pruned}.  Each method
  performs one evaluation on each of 240 processors per time step.  All
  algorithms use the Ring topology.}
  \label{fig:griewank-manyiters}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewankall}
  \caption{Function Griewank with 20 dimensions, comparing the performance of
  standard and speculative parallelizations as the number of processors
  increases.  All algorithms use the Ring topology.}
  \label{fig:griewank-all}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{bohachevsky}
  \caption{Function Bohachevsky with 20 dimensions.  Each method performs one
  evaluation on each of 480 processors per time step.}
  \label{fig:bohachevsky}
\end{figure}

\begin{table}
  \caption{Summary of results for function Bohachevsky with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06.}
  \label{tab:bohachevsky-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Random&100\%&472.3&6.6\\
  \hline
  Standard Complete&100\%&238.2&11.2\\
  \hline
  SEPSO Random&75\%&261.6&9.2\\
  \hline
  Pick Best Random&100\%&211.8&9.2\\
  \hline
  Pick Best Pruned Random&100\%&189.7&3.4\\
  \hline
  Pick Best Pruned Complete&25\%&105.4&3.9\\
  \hline
  Social Promotion Pruned Random&100\%&260.3&6.5\\
  \hline
  Social Promotion Pruned Complete&90\%&197.9&12.2\\
  \hline
  \textbf{Many Iterations Random}&\textbf{94\%}&\textbf{118.2}&\textbf{2.6}\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{Summary of results for function Rastrigin with 50 dimensions,
  measuring number of time steps to reach a value of 100}
  \label{tab:rastrigin-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&5\%&776.0&0.0\\
  \hline
  \textbf{Standard Random}&\textbf{100\%}&\textbf{593.2}&\textbf{89.5}\\
  \hline
  Standard Complete&80\%&343.8&145.8\\
  \hline
  Many Iterations Ring&25\%&223.8&67.1\\
  \hline
  Many Iterations Random&25\%&98.6&11.0\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Sphere with 50 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:sphere-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&30\%&986.8&8.2\\
  \hline
  Standard Complete&100\%&458.4&12.3\\
  \hline
  Many Iterations Ring&100\%&288.7&3.9\\
  \hline
  \textbf{Many Iterations Random}&\textbf{100\%}&\textbf{248.6}&\textbf{5.1}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Schwefel with 50 dimensions,
  measuring number of time steps to reach a value of 80}
  \label{tab:schwefel-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&70\%&871.6&57.1\\
  \hline
  Standard Complete&100\%&599.0&85.6\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{100\%}&\textbf{238.6}&\textbf{34.2}\\
  \hline
  \textbf{Many Iterations Random}&\textbf{100\%}&\textbf{242.9}&\textbf{95.4}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Griewank with 50 dimensions,
  measuring number of time steps to reach a value of 1e-06}
  \label{tab:griewank-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&1898.5&25.7\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{97\%}&\textbf{312.4}&\textbf{6.3}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Bohachevsky with 50 dimensions,
  measuring number of time steps to reach a value of 0.01}
  \label{tab:bohachevsky-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&1434.0&62.1\\
  \hline
  Standard Random&100\%&704.5&19.7\\
  \hline
  Standard Complete&20\%&377.5&14.9\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{95\%}&\textbf{229.8}&\textbf{11.4}\\
  \hline
  Many Iterations Random&35\%&171.6&4.0\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{rbf_stuck}
  \caption{Results for fitting a radial basis function to noisy data.  We use
  144 processors for each method, so each time step corresponds to 144 function
  evaluations.}
  \label{fig:rbf-stuck}
\end{figure}

>>>>

\begin{table}
  \caption{Summary of results for function Sphere with 20 dimensions,
  measuring number of time steps to reach a value of 1e-35.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:sphere-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&0\%&N/A&N/A\\
  \hline
  Standard Complete&100\%&931.0&19.1\\
  \hline
  SEPSO Ring&0\%&N/A&N/A\\
  \hline
  SEPSO Random&85\%&972.8&15.8\\
  \hline
  Pick Best Ring&100\%&917.4&21.0\\
  \hline
  Pick Best Random&100\%&768.1&14.3\\
  \hline
  Pick Best Pruned Ring&100\%&967.8&9.9\\
  \hline
  Pick Best Pruned Random&100\%&693.5&10.4\\
  \hline
  \textbf{Pick Best Pruned Complete}&\textbf{100\%}&\textbf{389.6}&
    \textbf{9.6}\\
  \hline
  Social Promotion Pruned Ring&0\%&N/A&N/A\\
  \hline
  Social Promotion Pruned Random&100\%&971.1&10.6\\
  \hline
  Social Promotion Pruned Complete&100\%&777.8&16.7\\
  \hline
  Many Iterations Ring&100\%&575.2&9.8\\
  \hline
  Many Iterations Random&100\%&442.4&7.1\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{Summary of results for function Schwefel with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:schwefel-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&0\%&N/A&N/A\\
  \hline
  Standard Complete&100\%&837.2&44.8\\
  \hline
  SEPSO Random&0\%&N/A&N/A\\
  \hline
  Pick Best Random&5\%&938.0&0.0\\
  \hline
  \textbf{Pick Best Pruned Random}&\textbf{100\%}&\textbf{639.9}&\textbf{29.4}\\
  \hline
  \textbf{Pick Best Pruned Complete}&\textbf{100\%}&\textbf{597.7}&
	\textbf{77.9}\\
  \hline
  Social Promotion Pruned Random&75\%&938.0&33.0\\
  \hline
  Social Promotion Pruned Complete&100\%&815.2&43.5\\
  \hline
  Many Iterations Random&100\%&783.5&94.5\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{Summary of results for function Rastrigin with 20 dimensions,
  measuring number of time steps to reach a value of 20.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:rastrigin-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&25\%&743.4&212.0\\
  \hline
  Standard Random&100\%&372.1&86.6\\
  \hline
  \textbf{Standard Complete}&\textbf{95\%}&\textbf{273.1}&\textbf{99.7}\\
  \hline
  SEPSO Ring&5\%&471.0&0.0\\
  \hline
  SEPSO Random&15\%&208.0&17.0\\
  \hline
  Pick Best Ring&35\%&489.0&152.0\\
  \hline
  Pick Best Random&15\%&136.0&40.1\\
  \hline
  Pick Best Pruned Ring&95\%&545.9&208.5\\
  \hline
  Pick Best Pruned Random&10\%&95.0&11.0\\
  \hline
  Pick Best Pruned Complete&5\%&59.0&0.0\\
  \hline
  Social Promotion Pruned Ring&5\%&580.0&0.0\\
  \hline
  Social Promotion Pruned Random&80\%&278.2&136.3\\
  \hline
  Social Promotion Pruned Complete&50\%&200.5&130.2\\
  \hline
  Many Iterations Ring&10\%&117.0&25.0\\
  \hline
  Many Iterations Random&5\%&57.0&0.0\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank6}
  \caption{Function Griewank with 20 dimensions.  Each method performs one
  evaluation on each of 800 processors per time step.  Instead of showing
  average function value, we show the percent of runs that are sufficiently
  close to the global optimum by each iteration ($10^{-6}$, as noted on the
  y-axis and explained in the text).  All algorithms use the Ring topology.}
  \label{fig:basic-griewank1}
\end{figure}

\begin{table}
  \caption{Summary of results for function Griewank with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:griewank-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&762.2&19.4\\
  \hline
  SEPSO Ring&100\%&426.0&81.0\\
  \hline
  \textbf{Pick Best Ring}&\textbf{100\%}&\textbf{272.0}&\textbf{17.7}\\
  \hline
  \textbf{Pick Best Pruned Ring}&\textbf{100\%}&\textbf{282.9}&\textbf{10.7}\\
  \hline
  Social Promotion Pruned Ring&100\%&482.0&16.6\\
  \hline
  Many Iterations Ring&84\%&183.2&49.4\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank1}
  \caption{Function Griewank with 20 dimensions.  Each method performs one
  evaluation of the objective function on each of 240 processors per time step.
  We show the percent of runs that are sufficiently close to the global optimum
  at each iteration.  All algorithms use the Ring topology.}
  \label{fig:basic-griewank2}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank3}
  \caption{Function Griewank with 20 dimensions, comparing Pick Best with
  results from \fig{basic-griewank2}.  Each method performs one evaluation on
  each of 240 processors per time step.  We show the percent of runs that are
  sufficiently close to the global optimum at each iteration.  All algorithms
  use the Ring topology.}
  \label{fig:griewank-pickbest}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank4}
  \caption{Function Griewank with 20 dimensions, comparing pruning with the
  best results from \fig{griewank-pickbest}.  Each method performs one
  evaluation on each of 240 processors per time step.  We show the percent of
  runs that are sufficiently close to the global optimum at each iteration.
  All algorithms use the Ring topology.}
  \label{fig:griewank-pruned}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewank5}
  \caption{Function Griewank with 20 dimensions, comparing speculating several
  iterations ahead with selections from \fig{griewank-pruned}.  Each method
  performs one evaluation on each of 240 processors per time step.  We show the
  percent of runs that are sufficiently close to the global optimum at each
  iteration.  All algorithms use the Ring topology.}
  \label{fig:griewank-manyiters}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{griewankall}
  \caption{Function Griewank with 20 dimensions, comparing the performance of
  standard and speculative parallelizations as the number of processors
  increases.  We show the percent of runs that are sufficiently close to the
  global optimum at each iteration.  All algorithms use the Ring topology.}
  \label{fig:griewank-all}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{bohachevsky}
  \caption{Function Bohachevsky with 20 dimensions.  Each method performs one
  evaluation on each of 480 processors per time step.  We show the percent of
  runs that are sufficiently close to the global optimum at each iteration.}
  \label{fig:bohachevsky}
\end{figure}

\begin{table}
  \caption{Summary of results for function Bohachevsky with 20 dimensions,
  measuring number of time steps to reach a value of 1e-06.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:bohachevsky-20}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Random&100\%&472.3&6.6\\
  \hline
  Standard Complete&100\%&238.2&11.2\\
  \hline
  SEPSO Random&75\%&261.6&9.2\\
  \hline
  Pick Best Random&100\%&211.8&9.2\\
  \hline
  Pick Best Pruned Random&100\%&189.7&3.4\\
  \hline
  Pick Best Pruned Complete&25\%&105.4&3.9\\
  \hline
  Social Promotion Pruned Random&100\%&260.3&6.5\\
  \hline
  Social Promotion Pruned Complete&90\%&197.9&12.2\\
  \hline
  \textbf{Many Iterations Random}&\textbf{94\%}&\textbf{118.2}&\textbf{2.6}\\
  \hline
  Many Iterations Complete&0\%&N/A&N/A\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{Summary of results for function Rastrigin with 50 dimensions,
  measuring number of time steps to reach a value of 100.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:rastrigin-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&5\%&776.0&0.0\\
  \hline
  \textbf{Standard Random}&\textbf{100\%}&\textbf{593.2}&\textbf{89.5}\\
  \hline
  Standard Complete&80\%&343.8&145.8\\
  \hline
  Many Iterations Ring&25\%&223.8&67.1\\
  \hline
  Many Iterations Random&25\%&98.6&11.0\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Sphere with 50 dimensions,
  measuring number of time steps to reach a value of 1e-06.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:sphere-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&30\%&986.8&8.2\\
  \hline
  Standard Complete&100\%&458.4&12.3\\
  \hline
  Many Iterations Ring&100\%&288.7&3.9\\
  \hline
  \textbf{Many Iterations Random}&\textbf{100\%}&\textbf{248.6}&\textbf{5.1}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Schwefel with 50 dimensions,
  measuring number of time steps to reach a value of 80.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:schwefel-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&0\%&N/A&N/A\\
  \hline
  Standard Random&70\%&871.6&57.1\\
  \hline
  Standard Complete&100\%&599.0&85.6\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{100\%}&\textbf{238.6}&\textbf{34.2}\\
  \hline
  \textbf{Many Iterations Random}&\textbf{100\%}&\textbf{242.9}&\textbf{95.4}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Griewank with 50 dimensions,
  measuring number of time steps to reach a value of 1e-06.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:griewank-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&1898.5&25.7\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{97\%}&\textbf{312.4}&\textbf{6.3}\\
  \hline
  \end{tabular}
\end{table}

\begin{table}
  \caption{Summary of results for function Bohachevsky with 50 dimensions,
  measuring number of time steps to reach a value of 0.01.  The ``best''
  algorithm is bolded, meaning that it has at least 90\% complete and its mean
  time to completion is significantly lower than all non-bolded algorithms,
  with a p-value of less than $10^{-5}$.}
  \label{tab:bohachevsky-50}
  \centering
  \begin{tabular}{|l|r|r|r|}
  \hline
  Algorithm&\% Complete&Mean&St. Dev.\\
  \hline
  \hline
  Standard Ring&100\%&1434.0&62.1\\
  \hline
  Standard Random&100\%&704.5&19.7\\
  \hline
  Standard Complete&20\%&377.5&14.9\\
  \hline
  \textbf{Many Iterations Ring}&\textbf{95\%}&\textbf{229.8}&\textbf{11.4}\\
  \hline
  Many Iterations Random&35\%&171.6&4.0\\
  \hline
  \end{tabular}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\columnwidth]{rbf_stuck}
  \caption{Results for fitting a radial basis function to noisy data.  We use
  144 processors for each method, so each time step corresponds to 144 function
  evaluations.  Here we show the percent of runs with a function value below
  55000, which we considered successful.}
  \label{fig:rbf-stuck}
\end{figure}
